---
title: "Derive growth parameters and natural mortality rates for krill while accounting for spatial heterogeneity in the Western Antarctic Peninsula."
subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
author: "Mardones, M; Cárdenas, C., Krüger, L., Santa Cruz, F."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: param.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup1}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.width = 8,
                      fig.height = 6,
                      dev = 'jpeg',
                      dpi = 300)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```



```{r lib, message=F, echo= TRUE}
library(here)
library(kableExtra)
library(ggthemes)
library(ggrepel)
library(ggridges)
#analisis
library(ggpubr)
library(easystats) # multiples unciones analiticas
library(sf)
library(tidyverse, quietly = TRUE)
library(modelsummary)
library(terra) # replace raster
library(TropFishR)
library(mixR)
library(CCAMLRGIS)
```


# Methodology

The object `ohbio2` come from data exploration analysis in data request
CCAMLR data. This objetc have bio information from krill.

```{r  echo=FALSE}
#datos entregados por secretaria de CCMLAR
metadata <- load("~/DOCAS/Data/565_C1_KRI_2021-10-01/DATA_PRODUCT_565.RData")
# Data procesada por MMardones
#ohbio <- load("DataLengthKrill.RData")
#ohbio
#metadata
#son lo mismo
#cargo objeto
meta <- get("METADATA")
c1 <- get("C1")
ohbio <- get("OBS_HAUL_BIOLOGY")
```

Join data set with master as `c1` set. This join is trought
`obs_haul_id` variable to get geoposition variables

```{r warning=FALSE}
ohbio2 <- left_join(c1, ohbio, by="obs_haul_id")
names(ohbio2)
```


```{r}
ohbio3 <- ohbio2 %>%
  mutate(Year = year(date_catchperiod_start),
         Month = month(date_catchperiod_start),
         Day = day(date_catchperiod_start)) %>% 
  #toupper() para convertir los valores a mayúsculas
  mutate(sex_code = toupper(sex_code)) %>% 
  dplyr::select(7, 9, 11, 12, 14, 24, 25, 29, 42, 44, 46, 47, 43, 45) |>  
  filter(asd_code=="481")
```


First thing is get different rater layer to join krill data length
according different porpoises.

```{r raster}
# Cargo linea de costa
coast <- load_Coastline()
coast1<- st_as_sf(coast) 
coast2 = st_transform(coast1, "+proj=latlong +ellps=WGS84")
# Uso las agrupaciones de Strata
strata <- st_read("~/DOCAS/Mapas/Antarctic_SHPfiles/Strata.shp",
                quiet=T)
strata=st_transform(strata, "+proj=latlong +ellps=WGS84")
```
Test Strata SSMU, just to know another way to join data, but this kind of spatial structuration is deprecated to mamagemente use (Figure\@ref(fig:maptest2).



Grouping bio data into stratas

```{r ssmu1}
ohbio6 <- st_as_sf(ohbio3 %>% 
                     drop_na(latitude_set_end), 
                   coords = c("longitude_set_end", "latitude_set_end"),  
                  crs = "+proj=latlong +ellps=WGS84")
```

## Study Area

```{r maptest, fig.cap="Strata Maps in 48.1"}
# y testeo el mapa
ssmap <- ggplot()+
  geom_sf(data = strata |> 
            filter(ID != "Outer") |>  
           mutate(ID = str_replace_all(ID, "Extra", "GS")),color="red")+
  geom_sf(data = ohbio6 |>
            drop_na(length_total_cm) |> 
            filter(Year>2017), 
          aes(color = length_total_cm)) +
  geom_text_repel(data = strata |>  
            filter(ID != "Outer") |>  
           mutate(ID = str_replace_all(ID, "Extra", "GS")), 
            aes(x = Labx, y = Laby, 
                label = ID), 
            min.segment.length = 0,
                             box.padding = 2,
                             max.overlaps = 10)+
  geom_sf(data = coast2, colour="black", fill=NA)+
  scale_color_viridis_c(option = "G",
                        name="Length (cm)")+
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  # coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)+
  theme_bw()
ssmap
```


Length composition by Strata CCAMLR to visualization first. First step is group data into to poligons strata.

```{r eval=FALSE}
strata <- st_make_valid(strata)
sf4 <- st_join(strata, ohbio6)
# Save an object to a file
saveRDS(sf4, file = "sf4.rds")

```

Load RData
```{r}
# Restore the object
sf4 <- readRDS("sf4.rds")
```

Statistical difference

```{r}
m <- aov(length_total_cm ~ID, data=sf4)
plot(TukeyHSD(m))
```


## Data

histogram length data to viz in anthor way.


```{r warning=FALSE}
jzstrata <- ggplot(sf4 %>% 
                     mutate(ID = if_else(ID == "Extra", "GERLASHE", ID)) %>% 
               filter(Year>2010,
                      ID !="Outer"),
             aes(x=length_total_cm, 
                 y = as.factor(Month), 
                 fill=ID))+
  geom_density_ridges(stat = "binline", bins = 30, 
                      scale = 1.9, 
                      draw_baseline = FALSE,
                      alpha=0.9)+
  facet_grid(Year~ID) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 2))+
  scale_y_discrete(breaks = seq(from = 1, 
                                to = 12, by = 4))+
  scale_fill_viridis_d(name="Strata",
                       option="F")+
  theme_few()+
  xlab("Length (cm.)")+
  ylab("")
jzstrata
```
Now, must filter the DF

```{r}
sf5 <- sf4 |> 
  dplyr::select(c("ID",
        "date_catchperiod_start",
        "length_total_cm",
        "Year",
        "Month",
        "sex_code",
         "greenweight_kg")) |> 
  mutate(ID = if_else(ID == "Extra", "GERLASHE", ID)) |>  
  filter(ID !="Outer") |> 
  data.frame()
```



Calculating the proportion of records per year and per stratum in a table. This table allows us to establish criteria for defining the strata on which we will analyze krill parameter. We leave out just `JOIN`.


```{r}
# Calcular la proporción
proporcion <- (round(table(sf5$ID) / sum(table(sf5$ID))*100,2))

# Crear una tabla con las proporciones
tabla_proporcion <- as.data.frame(proporcion)
tabla_proporcion$Categoria <- rownames(tabla_proporcion)
rownames(tabla_proporcion) <- NULL

# Renombrar las columnas
colnames(tabla_proporcion) <- c("Strata", "%")

# Mostrar la tabla utilizando kbl()
kbl(tabla_proporcion, 
    caption = "Proportion length register by Strata")  |> 
  kable_classic(full_width = F, 
                html_font = "Cambria") |> 
  kable_styling(bootstrap_options = "striped", 
                latex_options = "striped")
```

## Parameters estimation `k` and `L_inf_`



```{r}
#MALE

sf5fil <- sf5 |>
  filter(ID %in% c("BS", "EI", "GERLASHE", "SSIW"),
         sex_code %in% "M") |>  
  mutate(date_catchperiod_start = as.Date(date_catchperiod_start)) |>  
  mutate(yearly_Group = floor_date(date_catchperiod_start, "year")) |>   # New column
  drop_na(length_total_cm) 

# Definir los nombres correspondientes a cada objeto lfq_results
nombres <- c("SSI", "BS", "GERLASHE", "EI")
# Crear una lista para almacenar los resultados de cada ID
lfq_results <- list()
# Iterar sobre cada ID
for (id in unique(sf5fil$ID)) {
  # Obtener el nombre correspondiente
  nombre <- nombres[which(unique(sf5fil$ID) == id)]
  # Filtrar los datos para el ID actual
  df_id <- sf5fil %>% filter(ID == id)
  # Crear un objeto lfq para el ID actual
  lfq <- lfqCreate(data = df_id,
                   Lname = "length_total_cm",
                   Dname = "yearly_Group",
                   bin_size = 0.1)
  # Agregar el resultado a la lista
  lfq_results[[id]] <- lfq
  # Graficar el objeto lfq
  plot(lfq, Fname = "catch",
       main = nombre)
}

#FEMALE

sf5filhe <- sf5 |>
  filter(ID %in% c("BS", "EI", "GERLASHE", "SSIW"),
         sex_code %in% "F") |>  
  mutate(date_catchperiod_start = as.Date(date_catchperiod_start)) |>  
  mutate(yearly_Group = floor_date(date_catchperiod_start, "year")) |>   # New column
  drop_na(length_total_cm) 

# Definir los nombres correspondientes a cada objeto lfq_results
nombres <- c("SSI", "BS", "GERLASHE", "EI")
# Crear una lista para almacenar los resultados de cada ID
lfq_resultshe <- list()
# Iterar sobre cada ID
for (id in unique(sf5filhe$ID)) {
  # Obtener el nombre correspondiente
  nombre <- nombres[which(unique(sf5filhe$ID) == id)]
  # Filtrar los datos para el ID actual
  df_idhe <- sf5filhe %>% filter(ID == id)
  # Crear un objeto lfq para el ID actual
  lfqhe <- lfqCreate(data = df_idhe,
                   Lname = "length_total_cm",
                   Dname = "yearly_Group",
                   bin_size = 0.1)
  # Agregar el resultado a la lista
  lfq_resultshe[[id]] <- lfqhe
  # Graficar el objeto lfq
  plot(lfqhe, Fname = "catch",
       main = nombre)
}
```

Identified difference in length distribution just to male:

```{r message=FALSE}
Bhattacharya(lfq_results$EI)
```


```{r message=FALSE}
Bhattacharya(lfq_results$SSIW)
```


```{r message=FALSE}
Bhattacharya(lfq_results$BS)
```


```{r message=FALSE}
Bhattacharya(lfq_results$GERLASHE)
```

now we assign objet to male `lfq_result` and female `lfq_resulthe` and plot with `lfqRestructure()`


Male

```{r}
# Definir los nombres correspondientes a cada objeto lfq_results
nombres <- c("SSI", "BS", "GERLASHE", "EI")
# Iterar sobre cada objeto lfq almacenado en lfq_results
for (i in seq_along(lfq_results)) {
  lfq <- lfq_results[[i]]
  # Obtener el nombre correspondiente
  nombre <- nombres[i]
  # Restructurar el objeto lfq
  lfqbin <- lfqRestructure(lfq, MA = 3, addl.sqrt = TRUE)
  # Graficar el objeto lfq reestructurado
  plot(lfqbin, hist.col = c("white", "black"),
       image.col = c(rep(rgb(1,0.8,0.8),1000), "white", 
                     rep(rgb(0.8,0.8,1),1000)),
       ylim = c(0,max(lfqbin$midLengths+0.5)),
       main = nombre)
  # Ajustar curvas al objeto lfq reestructurado
  tmp <- lfqFitCurves(lfqbin, par = list(Linf=6.5, 
                                          K=0.45,
                                          t_anchor=0.5),
                      draw = TRUE, col=4, lty=2)
}
```

Female

```{r}
# Crear una lista para almacenar los resultados de PW_results <- list()

# Definir los nombres correspondientes a cada objeto lfq_results
nombres <- c("SSI", "BS", "GERLASHE", "EI")

# Iterar sobre cada objeto lfq almacenado en lfq_results
for (i in seq_along(lfq_resultshe)) {
  lfqhe <- lfq_resultshe[[i]]
  
  # Obtener el nombre correspondiente
  nombre <- nombres[i]
  
  # Restructurar el objeto lfq
  lfqbinhe <- lfqRestructure(lfqhe, MA = 3, addl.sqrt = TRUE)
  
  # Graficar el objeto lfq reestructurado
  plot(lfqbinhe, hist.col = c("white", "black"),
       image.col = c(rep(rgb(1,0.8,0.8),1000), "white", 
                     rep(rgb(0.8,0.8,1),1000)),
       ylim = c(0,max(lfqbinhe$midLengths+0.5)),
       main = nombre)
  
  # Ajustar curvas al objeto lfq reestructurado
  tmp <- lfqFitCurves(lfqbinhe, par = list(Linf=6.5, 
                                          K=0.45,
                                          t_anchor=0.5),
                      draw = TRUE, col=4, lty=2)
}

```
### Modal Progression analysis with `ELEFAN` to estimated `Linf`and `K`

Method based on @Pauly1987 and @Mildenberger2017

#### Krill Male

first we use `mixR` to identified numbers of compoenent (modal compositions) we have each stratum. With this result, set `ELEFAN` method


```{r eval=FALSE}
unique_ids <- unique(sf5$ID)

obj_id <- list()

for (id in unique_ids) {
  filtered_data <- sf5 %>% 
    drop_na(length_total_cm) %>% 
    filter(sex_code %in% "M", 
           ID !="JOIN",
           ID == id)
  obj_id[[id]] <- filtered_data
}

# select number of components

s_normalgs = select(obj_id$GERLASHE$length_total_cm, ncomp = 2:6)
plot(s_normalgs)
s_normalei = select(obj_id$EI$length_total_cm, ncomp = 2:6)
plot(s_normalei)
s_normalbs = select(obj_id$BS$length_total_cm, ncomp = 2:6)
plot(s_normalbs)
s_normalsswi = select(obj_id$SSWI$length_total_cm, ncomp = 2:6)
plot(s_normalsswi)

# fit a Normal mixture model
modgs = mixfit(obj_id$GERLASHE$length_total_cm, ncomp = 3,
              pi = c(0.5, 0.5, 0.5), 
              mu = c( 3, 4, 5), 
              sd = c(0.2, 0.2, 0.2))


# fit a Normal mixture model
modei = mixfit(obj_id$EI$length_total_cm, ncomp = 4,
              pi = c(0.5, 0.5, 0.5, 0.5), 
              mu = c( 3, 4, 5, 6), 
              sd = c(0.2, 0.2, 0.2, 0.2))


# fit a Normal mixture model
modbs = mixfit(obj_id$BS$length_total_cm, ncomp = 5,
              pi = c(0.5, 0.5, 0.5, 0.5, 0.5), 
              mu = c( 2, 3, 4, 5, 6), 
              sd = c(0.2, 0.2, 0.2, 0.2, 0.2))


# fit a Normal mixture model
modsswi = mixfit(obj_id$SSIW$length_total_cm, ncomp = 4,
              pi = c(0.5, 0.5, 0.5, 0.5), 
              mu = c( 3, 4, 5, 6), 
              sd = c(0.2, 0.2, 0.2, 0.2))
```


```{r, eval=FALSE, fig.width=8, fig.height=5}
GSplot <- plot(modgs,
     theme = "bw",
     title ="GERLASHE")
EIplot <- plot(modei,
     theme = "bw",
     title ="ELEFANT ISLAND")
BSplot <- plot(modbs,
     theme = "bw",
     title ="BRANSFIELD STRAIT")
SSWIplot <- plot(modsswi,
     theme = "bw",
     title ="SSWI")

ggarrange(GSplot, 
          EIplot, 
          BSplot, 
          SSWIplot, 
          common.legend = TRUE,
          ncol=4,
          legend="right")
```


```{r}
#### Parameters to `SSWI`

# run ELEFAN with simulated annealing
res_SAsswi <- ELEFAN_SA(lfq_results$SSIW, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAsswi <- ELEFAN_GA(lfq_results$SSIW, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                     low_par = list(Linf = 5,
                                   K = 0.5,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)

#### Parameters to `EI`

# run ELEFAN with simulated annealing
res_SAei <- ELEFAN_SA(lfq_results$EI, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAei <- ELEFAN_GA(lfq_results$EI, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                     low_par = list(Linf = 5,
                                   K = 0.5,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)


#### Parameters to `BS`

# run ELEFAN with simulated annealing
res_SAbs <- ELEFAN_SA(lfq_results$BS, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                   init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAbs <- ELEFAN_GA(lfq_results$BS, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)


#### Parameters to `GERLASHE`

# run ELEFAN with simulated annealing
res_SAgs <- ELEFAN_SA(lfq_results$GERLASHE, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAgs <- ELEFAN_GA(lfq_results$GERLASHE, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)
```
#### Krill Female


```{r}
#### Parameters to `SSWI`

# run ELEFAN with simulated annealing
res_SAsswihe <- ELEFAN_SA(lfq_resultshe$SSIW, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAsswihe <- ELEFAN_GA(lfq_resultshe$SSIW, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                     low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)


#### Parameters to `EI`

# run ELEFAN with simulated annealing
res_SAeihe <- ELEFAN_SA(lfq_resultshe$EI, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAeihe <- ELEFAN_GA(lfq_resultshe$EI, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                     low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)


#### Parameters to `BS`


# run ELEFAN with simulated annealing
res_SAbshe <- ELEFAN_SA(lfq_resultshe$BS, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                   init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAbshe <- ELEFAN_GA(lfq_resultshe$BS, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)


#### Parameters to `GERLASHE`


# run ELEFAN with simulated annealing
res_SAgshe <- ELEFAN_SA(lfq_resultshe$GERLASHE, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 6, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))

# run ELEFAN with genetic algorithm
res_GAgshe <- ELEFAN_GA(lfq_resultshe$GERLASHE, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 5,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 7.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)
```

Resuming parameters to both sex  by strata to krill
```{r}
#male
GAbs <- unlist(res_GAbs$par)
SAbs <- unlist(res_SAbs$par)
GAei <- unlist(res_GAei$par)
SAei <- unlist(res_SAei$par)
GAsswi <- unlist(res_GAsswi$par)
SAsswi <- unlist(res_SAsswi$par)
GAgs <- unlist(res_GAgs$par)
SAgs <- unlist(res_SAgs$par)
#female
GAbshe <- unlist(res_GAbshe$par)
SAbshe <- unlist(res_SAbshe$par)
GAeihe <- unlist(res_GAeihe$par)
SAeihe <- unlist(res_SAeihe$par)
GAsswihe <- unlist(res_GAsswihe$par)
SAsswihe <- unlist(res_SAsswihe$par)
GAgshe <- unlist(res_GAgshe$par)
SAgshe <- unlist(res_SAgshe$par)
# join 
t_k_linf_m <- rbind(GAbs[1:2],
                  SAbs[1:2],
                  GAei[1:2],
                  SAei[1:2],
                  GAsswi[1:2],
                  SAsswi[1:2],
                  GAgs[1:2],
                  SAgs[1:2])
t_k_linf_f <- rbind(GAbshe[1:2],
                  SAbshe[1:2],
                  GAeihe[1:2],
                  SAeihe[1:2],
                  GAsswihe[1:2],
                  SAsswihe[1:2],
                  GAgshe[1:2],
                  SAgshe[1:2])
row_names <- c("GA BS", "SA BS", "GA EI", 
                 "SA EI", "GA SSWI", "SA SSWI",
                 "GA GS", "SA GS")
rownames(t_k_linf_m) <- row_names
rownames(t_k_linf_f) <- row_names
col_names_m <- c("L inf Male", "K Male")
col_names_f <- c("L inf Female", "K Female")
colnames(t_k_linf_m) <- col_names_m
colnames(t_k_linf_f) <- col_names_f


total_para <- round(cbind(t_k_linf_f,
                    t_k_linf_m),3)
t_k_linf_total <- as_tibble(total_para, rownames = "group")


means <- t_k_linf_total %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
  mutate(group = "Mean")

t_k_linf <- bind_rows(t_k_linf_total, means)
```

```{r eval=FALSE}
write_csv(t_k_linf, "parametros_krill.csv")
```


total Table with `L_inf_` and `K`

```{r}
t_k_linf %>%
  separate(group, into = c("METHOD", "STRATA"), sep = " ") %>% 
  kbl(booktabs = T,
      position="ht!",
    caption = "Parametres LH in different strata with 2 algoritms to estimation") %>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE)

```

Made VB curves by methot and strata by male

separated group


```{r}
t_k_linf_sep <- t_k_linf %>% 
  filter(group != "Mean") %>% 
   separate(group, into = c("METHOD", "STRATA"), sep = " ")
```


```{r}
# Definir la función de Von Bertalanffy
von_bertalanffy <- function(t, L_inf, K) {
  L_inf * (1 - exp(-K * (t-t0)))
}

# Generar una secuencia de tiempos (por ejemplo, de 0 a 5 años)
t0 <- -0.3
time <- seq(0, 7, by = 0.1)

# Calcular longitud esperada para cada zona y tiempo
df_curves_male <- t_k_linf_sep %>%
  dplyr::select(c(1:2, 5:6)) %>% 
  rowwise() %>%
  mutate(length = list(von_bertalanffy(time, `L inf Male`, `K Male`))) %>%
  unnest(length)
# Agregar el vector de tiempo al DataFrame df_curves
df_curves_male$time <- rep(time, nrow(t_k_linf_sep))
# Graficar con ggplot
male <- ggplot(df_curves_male,
               aes(x = time, 
                              y = length,
                              group=STRATA,
                              colour=STRATA)) +
  geom_line(linewidth=1.1) +
  labs(title = "Male",
       x = "years",
       y = "Length (cm)") +
  theme_few()+
  facet_wrap(.~METHOD)+
  scale_colour_viridis_d(option="H",
                         name="Strata")

```
Made VB curves by methot and strata by female

```{r}
# Calcular longitud esperada para cada zona y tiempo
df_curves_female <- t_k_linf_sep %>%
  dplyr::select(c(1:4)) %>% 
  rowwise() %>%
  mutate(length = list(von_bertalanffy(time, `L inf Female`, `K Female`))) %>%
  unnest(length)
# Agregar el vector de tiempo al DataFrame df_curves
df_curves_female$time <- rep(time, nrow(t_k_linf_sep))
# Graficar con ggplot
female <- ggplot(df_curves_female,
               aes(x = time, 
                              y = length,
                              group=STRATA,
                              colour=STRATA)) +
  geom_line(linewidth=1.1) +
  labs(title = "Female",
       x = "years",
       y = "Length (cm)") +
  theme_few()+
  facet_wrap(.~METHOD)+
  scale_colour_viridis_d(option="H",
                         name="Strata")

```

both plot

```{r fig.width=10, fig.height=8}
ggarrange(female, male, common.legend = TRUE,
          ncol=1,
          legend="bottom")
```


### Method Nonlinear Mixed-Effects Models

Fit a nonlinear mixed-effects model (NLMM) to data, via maximum likelihood using `lme4` (@Bates2015) to calcularte parametrs

(work in progress)

```{r eval=FALSE, echo=TRUE}

#We use firts a descomposition methot with library `mixR` and the

sf52020 <- sf5 %>% 
  filter(sex_code %in% "F",
         ID == "EI",
         Year =="2020") %>% 
  drop_na(length_total_cm)

# fit a Normal mixture model
mod1 = mixfit(sf52020$length_total_cm, ncomp = 4,
              pi = c(0.5, 0.5, 0.5, 0.5), 
              mu = c( 3, 4, 5, 6), 
              sd = c(0.2, 0.2, 0.2, 0.2))
mod2_weibull = mixfit(sf6$length_total_cm, family = 'weibull', ncomp = 3)

s_weibull = select(sf5a$length_total_cm, ncomp = 2:6, family = 'weibull')

s_normal = select(sf5a$length_total_cm, ncomp = 2:6)
plot(s_weibull)
plot(s_normal)

# plot the fitted model# plot the fitted model# plot the fitted model
plot(mod1)
plot(mod2_weibull)

# fit a Normal mixture model (equal variance)
mod1_ev = mixfit(sf6$length_total_cm, ncomp = 2, ev = TRUE)

```

## Natural Mortality in Krill

In this exercises different bioanalogic methods are tested with General Algoritm

#### MALE

```{r}
MSSWI <- c(res_GAsswi$par, list(agemax = res_GAsswi$agemax))
# use the function M_empirical to estimate natural mortality
Msswi <- M_empirical(Linf = MSSWI$Linf, K_l = MSSWI$K, 
                  tmax = MSSWI$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                            "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# Bransfield
BS <- c(res_GAbs$par, list(agemax = res_GAbs$agemax))

# use the function M_empirical to estimate natural mortality
Mbs <- M_empirical(Linf = BS$Linf, K_l = BS$K, 
                  tmax = BS$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                              "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# GERLASHE
GS <- c(res_GAgs$par, list(agemax = res_GAgs$agemax))

# use the function M_empirical to estimate natural mortality
Mgs <- M_empirical(Linf = GS$Linf, K_l = GS$K, 
                  tmax = GS$agemax, temp = 5,
                   tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                             "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# Elephand Island
MEI <- c(res_GAei$par, list(agemax = res_GAei$agemax))

# use the function M_empirical to estimate natural mortality
Mei <- M_empirical(Linf = MEI$Linf, K_l = MEI$K, 
                  tmax = MEI$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                             "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
```

#### FEMALE

```{r}
MSSWIhe <- c(res_GAsswihe$par, list(agemax = res_GAsswihe$agemax))
# use the function M_empirical to estimate natural mortality
Msswihe <- M_empirical(Linf = MSSWIhe$Linf, K_l = MSSWIhe$K, 
                  tmax = MSSWIhe$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", 
                            "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# Bransfield
BShe <- c(res_GAbshe$par, list(agemax = res_GAbshe$agemax))

# use the function M_empirical to estimate natural mortality
Mbshe <- M_empirical(Linf = BShe$Linf, K_l = BShe$K, 
                  tmax = BShe$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", 
                              "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# GERLASHE
GShe <- c(res_GAgshe$par, list(agemax = res_GAgshe$agemax))

# use the function M_empirical to estimate natural mortality
Mgshe <- M_empirical(Linf = GShe$Linf, K_l = GShe$K, 
                  tmax = GShe$agemax, temp = 5,
                   tm50 = 2,
                  method = c("Pauly_Linf", 
                             "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# Elephand Island
MEIhe <- c(res_GAeihe$par, list(agemax = res_GAeihe$agemax))

# use the function M_empirical to estimate natural mortality
Meihe <- M_empirical(Linf = MEIhe$Linf, K_l = MEIhe$K, 
                  tmax = MEIhe$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", 
                             "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
```

```{r}
# junto las bases

Total_M_male <- cbind(Mei, Mbs, Mgs, Msswi)
colnames(Total_M_male) <- c("EI", "BS" , "GS", "SSWI")
mean_m <- colMeans(Total_M_male, na.rm = TRUE)
Total_M_Mean_male <- rbind(Total_M_male, Mean = mean_m)
# Agregar los promedios como una nueva fila al final del array

Total_M_femal <- cbind(Meihe, Mbshe, Mgshe, Msswihe)
colnames(Total_M_femal) <- c("EI", "BS" , "GS", "SSWI")
mean_m <- colMeans(Total_M_femal, na.rm = TRUE)
Total_M_Mean_femal <- rbind(Total_M_femal, Mean = mean_m)
```


```{r}
Total_M_Mean_male  %>%
  kbl(booktabs = T,
      position="ht!",
    caption = "Estimated M by Strata for Male") %>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE)

Total_M_Mean_femal  %>%
  kbl(booktabs = T,
      position="ht!",
    caption = "Estimated M by Strata for Female") %>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE)

```


Plot to female
```{r}
# Convertir la matriz a data.frame
df_female <- as.data.frame(Total_M_Mean_femal)

# Agregar la columna de métodos
df_female$Method <- rownames(Total_M_Mean_femal)

# Convertir el data.frame a formato largo
df_long <- df_female %>%
  pivot_longer(cols = EI:SSWI, 
               names_to = "Stratum", 
               values_to = "Value")

# Dot Plot
m_female<- ggplot(df_long %>%
                              drop_na() %>%
                              filter(Method != "Mean"), 
                            aes(x = Stratum, 
                                y = Value,
                                fill = Method)) +
    geom_point(size = 3,     
              shape = 21,   
              color = "black") + 
  geom_text_repel(aes(label = round(Value, 3)),  
                  size = 3,                      
                  box.padding = 0.35,            
                  point.padding = 0.5,         
                  segment.color = 'grey50',
                  min.segment.length = 0, 
                  nudge_y = 0.05,   
                  direction = "both") +
  labs(title = "Female",
       x = "",
       y = "Natural Mortality",
       fill = "Method") +
  theme_few() +
  scale_fill_viridis_d(option = "H") +
  ylim(0, 2)
```

Plot to male
```{r}
# Convertir la matriz a data.frame
df_male <- as.data.frame(Total_M_Mean_male)

# Agregar la columna de métodos
df_male$Method <- rownames(Total_M_Mean_male)

# Convertir el data.frame a formato largo
df_long_male <- df_male %>%
  pivot_longer(cols = EI:SSWI, 
               names_to = "Stratum", 
               values_to = "Value") 

# Dot Plot
m_male<- ggplot(df_long_male %>%
                              drop_na() %>%
                              filter(Method != "Mean"), 
                            aes(x = Stratum, 
                                y = Value,
                                fill = Method)) +
  geom_point(size = 3,     
              shape = 21,   
              color = "black") +  
  geom_text_repel(aes(label = round(Value, 3)),  
                  size = 3,                      
                  box.padding = 0.35,            
                  point.padding = 0.5,         
                  segment.color = 'grey50',
                  min.segment.length = 0, 
                  nudge_y = 0.05,   
                  direction = "both") +
  labs(title = "Male",
       x = "",
       y = "Natural Mortality",
       fill = "Method") +
  theme_few() +
  scale_fill_viridis_d(option = "H") +
  ylim(0, 2)

```

both plot

```{r}
ggarrange(m_female, m_male, 
          common.legend = TRUE,
          ncol=1,
          legend="right")
```
Statistical diferences in female

```{r}
# Realizar la prueba ANOVA
anova_result <- aov(Value ~ Stratum, data = df_long_male %>% drop_na())
summary(anova_result)
# Realizar la prueba post hoc de Tukey
tukey_result <- TukeyHSD(anova_result)
# Convertir los resultados de Tukey a un data frame
tukey_df <- as.data.frame(tukey_result$Stratum)
tukey_df$pair <- rownames(tukey_df)
tukey_df <- tukey_df %>%
  mutate(Stratum1 = sapply(strsplit(pair, "-"), `[`, 1),
         Stratum2 = sapply(strsplit(pair, "-"), `[`, 2))

# Seleccionar las columnas relevantes
tukey_table <- tukey_df %>%
  dplyr::select(Stratum1, Stratum2, diff, `p adj`, `lwr`, `upr`) %>%
  rename(Difference = diff, 
         P_value = `p adj`, 
         Lower_CI = `lwr`, 
         Upper_CI = `upr`)


kbl(tukey_table, 
    caption = "Test to differences between strata")  |> 
  kable_classic(full_width = F, 
                html_font = "Cambria") |> 
  kable_styling(bootstrap_options = "striped", 
                latex_options = "striped")



```
Statistical diferences in female

```{r}
# Realizar la prueba ANOVA
anova_result_ma <- aov(Value ~ Stratum, data = df_long %>% drop_na())
summary(anova_result_ma)
# Realizar la prueba post hoc de Tukey
tukey_result_ma <- TukeyHSD(anova_result_ma)
# Convertir los resultados de Tukey a un data frame
tukey_df_ma <- as.data.frame(tukey_result_ma$Stratum)
tukey_df_ma$pair <- rownames(tukey_df_ma)
tukey_df_ma <- tukey_df_ma %>%
  mutate(Stratum1 = sapply(strsplit(pair, "-"), `[`, 1),
         Stratum2 = sapply(strsplit(pair, "-"), `[`, 2))

# Seleccionar las columnas relevantes
tukey_table_ma <- tukey_df_ma %>%
  dplyr::select(Stratum1, Stratum2, diff, `p adj`, `lwr`, `upr`) %>%
  rename(Difference = diff, 
         P_value = `p adj`, 
         Lower_CI = `lwr`, 
         Upper_CI = `upr`)


kbl(tukey_table, 
    caption = "Test to differences between strata")  |> 
  kable_classic(full_width = F, 
                html_font = "Cambria") |> 
  kable_styling(bootstrap_options = "striped", 
                latex_options = "striped")


```


# References


