---
title: "Derive growth parameters and Natural Mortality in krill considering spatial heterogenity"
subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
author: "Mardones, M; Cárdenas, C."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: param.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup1}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.width = 8,
                      fig.height = 6,
                      dev = 'jpeg',
                      dpi = 300)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```



```{r lib, message=F, echo= TRUE}
library(here)
library(kableExtra)
library(ggthemes)
library(ggrepel)
library(ggridges)
#analisis
library(ggpubr)
library(easystats) # multiples unciones analiticas
library(sf)
library(tidyverse, quietly = TRUE)
library(modelsummary)
library(terra) # replace raster
library(TropFishR)
library(mixR)
library(CCAMLRGIS)
```


# Methodology

The object `ohbio2` come from data exploration analysis in data request
CCAMLR data. This objetc have bio information from krill.

```{r  echo=FALSE}
#datos entregados por secretaria de CCMLAR
metadata <- load("~/DOCAS/Data/565_C1_KRI_2021-10-01/DATA_PRODUCT_565.RData")
# Data procesada por MMardones
#ohbio <- load("DataLengthKrill.RData")
#ohbio
#metadata
#son lo mismo
```

```{r}
#cargo objeto
meta <- get("METADATA")
c1 <- get("C1")
ohbio <- get("OBS_HAUL_BIOLOGY")
```

Join data set with master as `c1` set. This join is trought
`obs_haul_id` variable to get geoposition variables

```{r warning=FALSE}
ohbio2 <- left_join(c1, ohbio, by="obs_haul_id")
names(ohbio2)
```

Firsts glance. Test how many register have by year. In this case,
`length_total_cm` by season ccamlr. Same exercise in date period
`date_catchperiod_start` to separate dates.

```{r}
ohbio3 <- ohbio2 %>%
  mutate(Year = year(date_catchperiod_start),
         Month = month(date_catchperiod_start),
         Day = day(date_catchperiod_start)) %>% 
  #toupper() para convertir los valores a mayúsculas
  mutate(sex_code = toupper(sex_code))
```

Save data further analysis
```{r}
length481 <-ohbio3 %>% 
  dplyr::select(7, 9, 11, 12, 14, 24, 25, 29, 42, 44, 46, 47, 43, 45) |>   filter(asd_code=="481")
#save(length481, file = "length481.RData")
```




First thing is get different rater layer to join krill data length
according different porpoises.

```{r raster}
# Cargo linea de costa
coast <- load_Coastline()
coast1<- st_as_sf(coast) 
coast2 = st_transform(coast1, "+proj=latlong +ellps=WGS84")
# Uso las agrupaciones de Strata
strata <- st_read("~/DOCAS/Mapas/Antarctic_SHPfiles/Strata.shp",
                quiet=T)
strata=st_transform(strata, "+proj=latlong +ellps=WGS84")
```
Test Strata SSMU, just to know another way to join data, but this kind of spatial structuration is deprecated to mamagemente use (Figure\@ref(fig:maptest2).



Grouping bio data into stratas

```{r ssmu1}
names(length481)
ohbio6 <- st_as_sf(length481 %>% 
                     drop_na(latitude_set_end), 
                   coords = c("longitude_set_end", "latitude_set_end"),  
                  crs = "+proj=latlong +ellps=WGS84")
```

## Study Area

```{r maptest, fig.cap="Strata Maps in 48.1"}
# y testeo el mapa
ssmap <- ggplot()+
  geom_sf(data = strata |> 
            filter(ID != "Outer") |>  
           mutate(ID = str_replace_all(ID, "Extra", "GS")),color="red")+
  geom_sf(data = ohbio6 |>
            drop_na(length_total_cm) |> 
            filter(Year>2017), 
          aes(color = length_total_cm)) +
  geom_text_repel(data = strata |>  
            filter(ID != "Outer") |>  
           mutate(ID = str_replace_all(ID, "Extra", "GS")), 
            aes(x = Labx, y = Laby, 
                label = ID), 
            min.segment.length = 0,
                             box.padding = 2,
                             max.overlaps = 10)+
  geom_sf(data = coast2, colour="black", fill=NA)+
  scale_color_viridis_c(option = "G",
                        name="Length (cm)")+
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  # coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)+
  theme_bw()
ssmap
```



Length composition by Strata CCAMLR to visualization first. First step is group data into to poligons strata.

```{r eval=FALSE}
strata <- st_make_valid(strata)
sf4 <- st_join(strata, ohbio6)
# Save an object to a file
saveRDS(sf4, file = "sf4.rds")

```

Load RData
```{r}
# Restore the object
sf4 <- readRDS("sf4.rds")
```



## Data

histogram length data to viz in anthor way.


```{r warning=FALSE}
jzstrata <- ggplot(sf4 %>% 
                     mutate(ID = if_else(ID == "Extra", "GERLASHE", ID)) %>% 
               filter(Year>2010,
                      ID !="Outer"),
             aes(x=length_total_cm, 
                 y = as.factor(Month), 
                 fill=ID))+
  geom_density_ridges(stat = "binline", bins = 30, 
                      scale = 1.9, 
                      draw_baseline = FALSE,
                      alpha=0.9)+
  facet_grid(Year~ID) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 2))+
  scale_y_discrete(breaks = seq(from = 1, 
                                to = 12, by = 4))+
  scale_fill_viridis_d(name="Strata",
                       option="F")+
  theme_few()+
  xlab("Length (cm.)")+
  ylab("")
jzstrata
```
Now, must filter the DF

```{r}
sf5 <- sf4 |> 
  dplyr::select(c("ID",
        "date_catchperiod_start",
        "length_total_cm",
        "Year",
        "Month",
        "sex_code",
         "greenweight_kg")) |> 
  mutate(ID = if_else(ID == "Extra", "GERLASHE", ID)) |>  
  filter(ID !="Outer") |> 
  data.frame()
```



Calculating the proportion of records per year and per stratum in a table. This table allows us to establish criteria for defining the strata on which we will analyze krill parameter. We leave out just `JOIN`.


```{r}
# Calcular la proporción
proporcion <- (round(table(sf5$ID) / sum(table(sf5$ID))*100,2))

# Crear una tabla con las proporciones
tabla_proporcion <- as.data.frame(proporcion)
tabla_proporcion$Categoria <- rownames(tabla_proporcion)
rownames(tabla_proporcion) <- NULL

# Renombrar las columnas
colnames(tabla_proporcion) <- c("Strata", "%")

# Mostrar la tabla utilizando kbl()
kbl(tabla_proporcion, 
    caption = "Proportion length register by Strata")  |> 
  kable_classic(full_width = F, 
                html_font = "Cambria") |> 
  kable_styling(bootstrap_options = "striped", 
                latex_options = "striped")
```

## Parameters estimation `k` and `L_inf_`

### Method `ELEFAN` 

(@Tobias2017)


```{r}
sf5fil <- sf5 |>
  filter(ID %in% c("BS", "EI", "GERLASHE", "SSIW")) |>  
  mutate(date_catchperiod_start = as.Date(date_catchperiod_start)) |>  
  mutate(yearly_Group = floor_date(date_catchperiod_start, "year")) |>   # New column
  drop_na(length_total_cm) 

# Definir los nombres correspondientes a cada objeto lfq_results
nombres <- c("SSI", "BS", "GERLASHE", "EI")
# Crear una lista para almacenar los resultados de cada ID
lfq_results <- list()
# Iterar sobre cada ID
for (id in unique(sf5fil$ID)) {
  # Obtener el nombre correspondiente
  nombre <- nombres[which(unique(sf5fil$ID) == id)]
  # Filtrar los datos para el ID actual
  df_id <- sf5fil %>% filter(ID == id)
  # Crear un objeto lfq para el ID actual
  lfq <- lfqCreate(data = df_id,
                   Lname = "length_total_cm",
                   Dname = "yearly_Group",
                   bin_size = 0.1)
  # Agregar el resultado a la lista
  lfq_results[[id]] <- lfq
  # Graficar el objeto lfq
  plot(lfq, Fname = "catch",
       main = nombre)
}

```
Identified distribution

```{r message=FALSE}

ei <- Bhattacharya(lfq_results$EI)
ss <- Bhattacharya(lfq_results$SSIW)
bs <- Bhattacharya(lfq_results$BS)
gs <- Bhattacharya(lfq_results$GERLASHE)

```
ahora damos asignaciones a curvas para representación

```{r}
# Crear una lista para almacenar los resultados de PW_results <- list()

# Definir los nombres correspondientes a cada objeto lfq_results
nombres <- c("SSI", "BS", "GERLASHE", "EI")

# Iterar sobre cada objeto lfq almacenado en lfq_results
for (i in seq_along(lfq_results)) {
  lfq <- lfq_results[[i]]
  
  # Obtener el nombre correspondiente
  nombre <- nombres[i]
  
  # Restructurar el objeto lfq
  lfqbin <- lfqRestructure(lfq, MA = 3, addl.sqrt = TRUE)
  
  # Graficar el objeto lfq reestructurado
  plot(lfqbin, hist.col = c("white", "black"),
       image.col = c(rep(rgb(1,0.8,0.8),1000), "white", 
                     rep(rgb(0.8,0.8,1),1000)),
       ylim = c(0,max(lfqbin$midLengths+0.5)),
       main = nombre)
  
  # Ajustar curvas al objeto lfq reestructurado
  tmp <- lfqFitCurves(lfqbin, par = list(Linf=6.5, 
                                          K=0.45,
                                          t_anchor=0.5),
                      draw = TRUE, col=4, lty=2)
}

```



Proceed to fit Modal Progression analysis with `ELEFAN`

#### Parameters to `SSIW`

```{r}
# Response surface analyss
res_RSAsswi <- ELEFAN(lfq = lfq_results$SSIW,  
                  MA = 5,
                  Linf_range = seq(4.5, 7, 
                                   length.out = 30),
                  K_range = exp(seq(log(0.1),
                                    log(2),
                                    length.out = 30)),
                  C = 0.5,
                  ts = 0.5,
                  method = "cross",
                  cross.date = lfq_results$SSIW$dates[3],
                  cross.midLength = lfq_results$SSIW$midLengths[5],
                  contour = TRUE,
                  add.values = FALSE,
                  agemax = 5,
                  hide.progressbar = TRUE)
```


```{r}
# run ELEFAN with simulated annealing
res_SAsswi <- ELEFAN_SA(lfq_results$SSIW, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 4.5, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 4,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))
```


```{r}
# run ELEFAN with genetic algorithm
res_GAsswi <- ELEFAN_GA(lfq_results$SSIW, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 4.5, 
                                   K = 0.01, 
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5, 
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)
```

Table with different method

```{r}
# show results
RSAsswi <- unlist(res_RSAsswi$par)
GAsswi <- unlist(res_GAsswi$par)
SAsswi <- unlist(res_SAsswi$par)
parsswi <- round(rbind(RSAsswi, GAsswi, SAsswi),3)

parsswi %>%
  kbl(booktabs = T,
      position="ht!",
    caption = "Parametres LH to SSWI") %>%
    kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE)

```
#### Parameters to `EI`

```{r}
# Response surface analyss
res_RSAei <- ELEFAN(lfq = lfq_results$EI,  
                  MA = 5,
                  Linf_range = seq(4.5, 7, 
                                   length.out = 30),
                  K_range = exp(seq(log(0.1),
                                    log(2),
                                    length.out = 30)),
                  C = 0.5,
                  ts = 0.5,
                  method = "cross",
                  cross.date = lfq_results$EI$dates[3],
                  cross.midLength = lfq_results$EI$midLengths[5],
                  contour = TRUE,
                  add.values = FALSE,
                  agemax = 5,
                  hide.progressbar = TRUE)
```


```{r}
# run ELEFAN with simulated annealing
res_SAei <- ELEFAN_SA(lfq_results$EI, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 4.5, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 4,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))
```


```{r}
# run ELEFAN with genetic algorithm
res_GAei <- ELEFAN_GA(lfq_results$EI, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 4.5, 
                                   K = 0.01, 
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5, 
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)
```

```{r}
# show results
RSAei <- unlist(res_RSAei$par)
GAei <- unlist(res_GAei$par)
SAei <- unlist(res_SAei$par)
parei <- round(rbind(RSAei, GAei, SAei),3)
```



#### Parameters to `BS`

```{r}
# Response surface analyss
res_RSAbs <- ELEFAN(lfq = lfq_results$BS,  
                  MA = 5,
                  Linf_range = seq(4.5, 7, 
                                   length.out = 30),
                  K_range = exp(seq(log(0.1),
                                    log(2),
                                    length.out = 30)),
                  C = 0.5,
                  ts = 0.5,
                  method = "cross",
                  cross.date = lfq_results$BS$dates[3],
                  cross.midLength = lfq_results$BS$midLengths[5],
                  contour = TRUE,
                  add.values = FALSE,
                  agemax = 5,
                  hide.progressbar = TRUE)
```


```{r}
# run ELEFAN with simulated annealing
res_SAbs <- ELEFAN_SA(lfq_results$BS, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 4.5, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 4,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))
```


```{r}
# run ELEFAN with genetic algorithm
res_GAbs <- ELEFAN_GA(lfq_results$BS, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 4.5, 
                                   K = 0.01, 
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5, 
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)
```

```{r}
# show results
RSAbs <- unlist(res_RSAbs$par)
GAbs <- unlist(res_GAbs$par)
SAbs <- unlist(res_SAbs$par)
parbs <- round(rbind(RSAbs, GAbs, SAbs),3)

```

#### Parameters to `GERLASHE`

```{r}
# Response surface analyss
res_RSAgs <- ELEFAN(lfq = lfq_results$GERLASHE,  
                  MA = 5,
                  Linf_range = seq(4.5, 7, 
                                   length.out = 30),
                  K_range = exp(seq(log(0.1),
                                    log(2),
                                    length.out = 30)),
                  C = 0.5,
                  ts = 0.5,
                  method = "cross",
                  cross.date = lfq_results$GERLASHE$dates[3],
                  cross.midLength = lfq_results$GERLASHE$midLengths[5],
                  contour = TRUE,
                  add.values = FALSE,
                  agemax = 5,
                  hide.progressbar = TRUE)
```


```{r}
# run ELEFAN with simulated annealing
res_SAgs <- ELEFAN_SA(lfq_results$GERLASHE, 
                    SA_time = 60*0.5, 
                    MA = 5, 
                    agemax = 5,
                    seasonalised = TRUE, addl.sqrt = TRUE,
                    init_par = list(Linf = 4.5, 
                                    K = 0.5, 
                                    t_anchor = 0.5,
                                    C=0.5, 
                                    ts = 0.5),
                    low_par = list(Linf = 4,
                                   K = 0.01,
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5,
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1))
```


```{r}
# run ELEFAN with genetic algorithm
res_GAgs <- ELEFAN_GA(lfq_results$GERLASHE, 
                    MA = 5, 
                    seasonalised = TRUE, 
                    maxiter = 10, 
                    agemax = 5,
                    addl.sqrt = TRUE,
                    low_par = list(Linf = 4.5, 
                                   K = 0.01, 
                                   t_anchor = 0, 
                                   C = 0, 
                                   ts = 0),
                    up_par = list(Linf = 6.5, 
                                  K = 1, 
                                  t_anchor = 1,
                                  C = 1, ts = 1),
                    monitor = FALSE)
```

```{r}
# show results
RSAgs <- unlist(res_RSAgs$par)
GAgs <- unlist(res_GAgs$par)
SAgs <- unlist(res_SAgs$par)
pargs <- round(rbind(RSAgs, GAgs, SAgs),3)

```
total Table with `L_inf_` and `K`

```{r}

t_k_linf <- rbind(parsswi, parei, parbs, pargs)
t_k_linf %>%
  kbl(booktabs = T,
      position="ht!",
    caption = "Parametres LH in different strata with 3 types of algoritms") %>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE)

```


### Method NO LINEAR MODEL 


We use firts a descomposition methot with library `mixR` and then we use `nlmer` (@Bates2015) to calcularte parametrs

```{r eval=FALSE, echo=TRUE}
# Agregar la nueva columna con los intervalos de longitud total
sf6 <- sf5 |> 
  drop_na(length_total_cm) |> 
  filter(ID %in% "GERLASHE",
         Year==2018,
         Month==5)

sf5a <-  sf5 |> 
  drop_na(length_total_cm ) |> 
  filter(ID %in% "BS",
         Year==2018)
  
# compruebo el n de comp

s_weibull = select(sf5a$length_total_cm, ncomp = 2:6, family = 'weibull')

s_normal = select(sf5a$length_total_cm, ncomp = 2:6)




# fit a Normal mixture model
mod1 = mixfit(sf5a$length_total_cm, ncomp = 4,
              pi = c(0.5, 0.5, 0.5, 0.5), 
              mu = c( 3, 4, 5, 6), 
              sd = c(0.2, 0.2, 0.2, 0.2))
mod2_weibull = mixfit(sf6$length_total_cm, family = 'weibull', ncomp = 3)

s_weibull = select(sf5a$length_total_cm, ncomp = 2:6, family = 'weibull')

s_normal = select(sf5a$length_total_cm, ncomp = 2:6)
plot(s_weibull)
plot(s_normal)

# plot the fitted model# plot the fitted model# plot the fitted model
plot(mod1)
plot(mod2_weibull)

# fit a Normal mixture model (equal variance)
mod1_ev = mixfit(sf6$length_total_cm, ncomp = 2, ev = TRUE)

```

## Natural Mortality in Krill

In this exercises different bioanalogic methods are tested with General Algoritm
```{r}
MSSWI <- c(res_GAsswi$par, list(agemax = res_RSAsswi$agemax))
# use the function M_empirical to estimate natural mortality
Msswi <- M_empirical(Linf = MSSWI$Linf, K_l = MSSWI$K, 
                  tmax = MSSWI$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                            "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# Bransfield
BS <- c(res_GAbs$par, list(agemax = res_RSAbs$agemax))

# use the function M_empirical to estimate natural mortality
Mbs <- M_empirical(Linf = BS$Linf, K_l = BS$K, 
                  tmax = BS$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                              "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# GERLASHE
GS <- c(res_GAgs$par, list(agemax = res_RSAgs$agemax))

# use the function M_empirical to estimate natural mortality
Mgs <- M_empirical(Linf = GS$Linf, K_l = GS$K, 
                  tmax = GS$agemax, temp = 5,
                   tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                             "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
# Elephand Island
MEI <- c(res_GAei$par, list(agemax = res_RSAei$agemax))

# use the function M_empirical to estimate natural mortality
Mei <- M_empirical(Linf = MEI$Linf, K_l = MEI$K, 
                  tmax = MEI$agemax, temp = 5,
                  tm50 = 2,
                  method = c("Pauly_Linf", "Hoenig", 
                             "Hoenig",
                             "AlversonCarney",
                             "RikhterEfanov"
                             ))
```

```{r}
# junto las bases

Total_M <- cbind(Mei, Mbs, Mgs, Msswi)
colnames(Total_M) <- c("EI", "BS" , "GS", "SSWI")
mean <- colMeans(Total_M, na.rm = TRUE)
# Agregar los promedios como una nueva fila al final del array
Total_M_Mean <- rbind(Total_M, Mean = mean)
```


```{r}
Total_M_Mean  %>%
  kbl(booktabs = T,
      position="ht!",
    caption = "Estimated M by Strata") %>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE)

```


# References


